---
title: "DATA557_FinalProj_Analysis"
output: html_document
---

```{r Load data}
data <- read.csv("/Users/nickw/OneDrive/Documents/School/DATA557/final_dataset.csv")
```

```{r Data Prep}
data$minutes_played <- gsub(",","",data$minutes_played)
data <- data[as.numeric(data$minutes_played) > 360,]
data$position <- substr(data$position,0,2)
data$league <- substr(data$league,0,2)
data$league <- tolower(data$league)
data$price_num <- gsub(',','', data$price_num)
data$price_num <- as.numeric(data$price_num)
data$price_num <- ifelse(tolower(data$price) == 'free', 0, data$price_num)
```

Exploratory Analysis
```{r Plot distributions}
# Distribution Plotting
hist(data$points_diff)
hist(data$overall_rating)
hist(as.numeric(data$minutes_played))
plot(data$overall_rating,data$points_diff, main = "Association between overall rating and point difference",xlab = "Overall Rating",ylab = "Point Difference")
plot(data$rank_before,data$points_diff, main = "Association between team ranking and point difference",xlab = "Team Ranking",ylab = "Point Difference")

# Stratified Correlation Analysis
x <- data[data$points_diff > 0,]
cor(x$points_diff,x$overall_rating)
x <- data[data$points_diff < 0,]
cor(x$points_diff,x$overall_rating)

# Transfer-Rating Correlation
cor(data$price_num,data$overall_rating)
plot(data$overall_rating,data$price_num)
```

Question 1

Question 1: How does a traded player's overall rating affect the success of the team that they were traded to (measured in point differential between the season before and the season after arrival)?
Outcome: difference in points in this season from the previous season
Primary predictor: the overall rating of the player before he was traded.
Potential confounders: player's position, team, season, league, number of minutes played

```{r Question 1}
summary(lm(points_diff ~ overall_rating+factor(position)+factor(season)+as.numeric(minutes_played)+rank_before,data = data))
```

Question 2:  Whether the position affect the association between success score and rating
Outcome: difference in success score in this season from the previous season
Primary predictor: the overall rating of the player before he was traded
Interaction term: position (4 positions) and rating
Potential confounders: Parameters that will be included in the model- player's position, season, league, number of minutes played

```{r Question 2}
summary(lm(points_diff ~ position:overall_rating+overall_rating+factor(position)+factor(season)+as.numeric(minutes_played)+rank_before, data=data))
```

Question 3: Does the rating affect the association between transfer money and the impact of each player. We will use a linear regression model. The outcome for this model is the effect of each player using the beta coefficient of the rating of the first model. 
Outcome: Transfer money
Primary predictor: player's impact- the difference in points in this season from the previous season times the coefficient of the rating variable in the first model
Interaction term: player's impact and rating
Potential confounders: player's position, season, league, number of minutes played

```{r Question 3}
data <- data[!is.na(data$price_num), ]
summary(lm(price_num ~ points_diff:overall_rating+points_diff+overall_rating+factor(position)+factor(season)+as.numeric(minutes_played)+rank_before, data=data))
```
